Full Dissertation

### 2024-04-12 Sections/hold_temp.qmd
re-render

>>>>

Write a complete paragraph for the "Existing Theoretical Frameworks" sections at the end of the Introduction. Cite papers like on Bjork, 1994; Soderstrom & Bjork, 2015; Guadagnoli & Lee, 2004, and make the point that while these existing frameworks offer a conceptual framework for understanding the influence of variability, there is currently a dearth of computational modelling approaches being utilized to explain the effects of varied training in the visuomotor learning domain.

### Existing Theoretical Frameworks

Several theoretical frameworks have been proposed to account for the effects of training variability on learning and generalization. The contextual interference effect (Battig, 1972; Shea & Morgan, 1979) posits that varied practice leads to greater interference during acquisition but enhanced retention and transfer due to increased contextual discrimination and cognitive effort during learning. Similarly, the reconceptualized desirable difficulties framework (Bjork, 1994; Soderstrom & Bjork, 2015) suggests that varied practice introduces desirable difficulties that initially impair performance but promote long-term retention and transfer. The challenge point hypothesis (Guadagnoli & Lee, 2004) proposes that optimal learning occurs when the difficulty of the task is matched to the learner's skill level, with varied practice serving to increase task difficulty and facilitate learning for more skilled individuals. While these frameworks offer conceptual explanations for the influence of variability, there is currently a paucity of computational modeling approaches that explicitly capture the mechanisms underlying the effects of varied training, particularly in the domain of visuomotor learning. Existing computational models have largely focused on category learning (e.g., Nosofsky, 1986) and perceptual learning (e.g., Petrov et al., 2005), with limited attention to the role of variability in visuomotor skill acquisition and generalization. Given the importance of understanding the cognitive processes involved in visuomotor learning and transfer, there is a need for computational models that can provide a more precise and quantitative account of how varied practice influences the acquisition and generalization of visuomotor skills.

Existing theoretical frameworks, such as the desirable difficulties framework (Bjork, 1994) and the challenge point framework (Guadagnoli & Lee, 2004), provide valuable conceptual insights into the influence of training variability on learning and generalization. These frameworks suggest that introducing appropriate levels of challenge or difficulty during training, such as through increased variability, can enhance long-term retention and transfer (Soderstrom & Bjork, 2015). However, while these frameworks offer a high-level understanding of the mechanisms underlying the benefits of varied training, there is currently a scarcity of computational modeling approaches being applied to elucidate the specific cognitive processes and representations involved in the effects of training variability, particularly in the domain of visuomotor learning. Computational models, such as connectionist or Bayesian approaches, have the potential to provide a more precise and quantitative account of how variability during training influences the acquisition and generalization of visuomotor skills. By specifying the underlying mechanisms and parameters, these models can generate testable predictions and offer a deeper understanding of the cognitive processes that give rise to the observed effects of varied training on visuomotor learning and transfer.

The influence of training variability on learning and generalization has been a topic of considerable interest within cognitive science, particularly in the context of motor learning and skill acquisition. Theoretical frameworks such as Bjork's (1994) "desirable difficulties" and Soderstrom and Bjork's (2015) elaboration on the role of varied practice highlight the potential benefits of introducing variability during training to enhance learning outcomes. These theories suggest that varying the conditions of practice can improve the adaptability and robustness of learned skills, a principle also supported by Guadagnoli and Lee's (2004) challenge point framework, which posits that optimal learning occurs when the difficulty of practice tasks is appropriately matched to the learner's capabilities. While these frameworks provide valuable insights into the mechanisms through which variability might enhance learning, they predominantly focus on general psychological and pedagogical principles. Notably, there is a relative scarcity of computational modeling approaches that explicitly address how varied training influences learning processes in the visuomotor domain. Such models are crucial for quantitatively testing theories and providing a more detailed mechanistic understanding of how variability affects the acquisition, retention, and generalization of motor skills. As such, the current research seeks to fill this gap by developing and applying computational models to empirically investigate the effects of training variability in visuomotor tasks, thereby providing a more nuanced understanding of the conditions under which variability facilitates or impedes learning and transfer.
